\textbf{ Mapping stellar surfaces\\ I: Degeneracies in the rotational light curve problem }

### 2021-01-27 paper1/ms.tex
finished implementing hogg's comments

The advent of space-based precision photometry with missions such as
\emph{Kepler} \citep{Borucki2010} and \emph{TESS} \citep{Ricker2015}
has led to a renewed interest in the modeling of stellar light curves,
and, in particular, in understanding what these light curves can tell
us about the surfaces of stars across the HR diagram. One of the dominant
sources of stellar light curve variability is the modulation caused
by starspots rotating in and out of view.
Dark spots arise due to the suppression of convection in regions of
intense magnetic field, resulting in a locally cooler (and hence darker)
photosphere. Bright spots can similarly
arise in the photosphere as faculae or in the chromosphere as plages, and
are also magnetically driven \citep[e.g.,][]{Berdyugina2005}.
%
Constraining the sizes,
contrasts, locations, and number of spots on stars can therefore reveal
information about stellar magnetic activity, stellar interior structure,
and how these quantities vary across spectral type and over time.
A detailed understanding of the stellar surface is also crucial to
mitigating systematics in the radial velocity search for exoplanets
\citep[e.g.,][]{Lanza2011} and in the spectroscopic characterization of their
atmospheres \citep[e.g.,][]{Rackham2018}.

@@

In this section, we define the concept of the \emph{null space}
and present some demonstrations showing how it can affect inferences
about stellar surfaces.
We also present some examples of how the degeneracies due to
the null space can be tackled.
While our specific examples may appear somewhat idealized, at the end
of this section we discuss how our ideas generalize to more
realistic scenarios.
This is also a topic that we discuss in much more detail in \citetalias{PaperII}.

@@

The second point is that while observing many similar stars at different
inclinations can greatly help us learn about their surfaces from a
statistical standpoint, our analysis
above assumed we knew what the values of the individual inclinations were.
In practice, this will usually not be the case. While we may have
good priors for some stars (from spectroscopic $v\sin i$ measurements,
or from the assumption that transiting exoplanet hosts are likely
to have inclinations close to $I=90^\circ$), for the vast majority
of field stars we won't know much a priori. Since the inclination
is typically degenerate with the spot latitude \citep[e.g.,][]{Walkowicz2013},
this decreases the constraining power of ensemble analyses.
However, as we show in \citetalias{PaperII}, there is still enough
information in a large ensemble to independently constrain the spot latitudes
and the \emph{individual} stellar inclinations.

@@

A direct consequence of this normalization degeneracy is that it is not
possible to uniquely constrain the total spot coverage of a star from
single-band photometry without strong prior assumptions.
The total spot coverage $f_S$
is simply the (average) area of a spot
times the total number of spots divided by the total area of the sphere, which
may be expressed as
%
\begin{proof}{test_fS}
    \label{eq:fS}
    f_S = \frac{1}{2}\left(1 - \left<\cos r\right>\right)n
    \quad,
\end{proof}
%
given an angular spot radius $r$.
While $r$ may be constrained from the covariance structure of the data
\citepalias{PaperII}, $n$ cannot.

@@

>>>>

The advent of space-based precision photometry with missions such as
\emph{Kepler} \citep{Borucki2010} and \emph{TESS} \citep{Ricker2015}
has led to a renewed interest in the modeling of stellar light curves,
and, in particular, in understanding what these light curves can tell
us about the surfaces of stars across the HR diagram. One of the dominant
sources of stellar light curve variability is the modulation caused
by starspots rotating in and out of view.
Dark spots arise due to the suppression of convection in regions of
intense magnetic field, resulting in a locally cooler (and hence darker)
photosphere. Bright spots can similarly
arise in the photosphere as faculae or in the chromosphere as plages, and
are also magnetically driven \citep[e.g.,][]{Berdyugina2005}.
%
Constraining the sizes,
contrasts, locations, and number of spots on stars can therefore reveal
information about stellar magnetic activity, stellar interior structure,
and how these quantities vary across spectral type and over time
\citep[e.g.,][]{Garraffo2018}.
A detailed understanding of the stellar surface is also crucial to
mitigating systematics in the radial velocity search for exoplanets
\citep[e.g.,][]{Lanza2011} and in the spectroscopic characterization of their
atmospheres \citep[e.g.,][]{Rackham2018}.

@@

In this section, we define the concept of the \emph{null space}
and present some demonstrations showing how it can affect inferences
about stellar surfaces.
For simplicity, we assume we know quantities like the stellar
inclination, rotation period, and limb darkening coefficients
exactly, and we assume the stellar surface does not vary in time.
Because of these assumptions, our constraints on the information
content of light curves should be viewed as strict \emph{upper limits}.

We also present some examples of how the degeneracies due to
the null space can be tackled.
While our examples below may appear somewhat idealized, at the end
of this section we discuss how our ideas generalize to more
realistic scenario; we also revisit these assumptions in \S\ref{sec:caveats}.
This is also a topic that we discuss in much more detail in \citetalias{PaperII}.

@@

The second point is that while observing many similar stars at different
inclinations can greatly help us learn about their surfaces from a
statistical standpoint, our analysis
above assumed we knew what the values of the individual inclinations were.
In practice, this will usually not be the case. While we may have
good priors for some stars (from spectroscopic $v\sin I$ measurements,
or from the assumption that transiting exoplanet hosts are likely
to have inclinations close to $I=90^\circ$), for the vast majority
of field stars we won't know much a priori. Since the inclination
is typically degenerate with the spot latitude \citep[e.g.,][]{Walkowicz2013},
this decreases the constraining power of ensemble analyses.
However, as we show in \citetalias{PaperII}, there is still enough
information in a large ensemble to independently constrain the spot latitudes
and the \emph{individual} stellar inclinations.

@@

A direct consequence of this normalization degeneracy is that it is not
possible to uniquely constrain the total spot coverage of a star from
single-band photometry without strong prior assumptions.
The total spot coverage $f_S$
is simply the (average) area of a spot
times the total number of spots divided by the total area of the sphere, which
may be expressed as
%
\begin{proof}{test_fS}
    \label{eq:fS}
    f_S = \frac{1}{2}\left(1 - \left<\cos r\right>\right)n
    \quad,
\end{proof}
%
given an angular spot radius $r$.
While $r$ may be uniquely constrained from the covariance structure of the data
\citepalias{PaperII}, $n$ cannot.

@@

\subsection{Caveats}
\label{sec:caveats}

There are several simplifying assumptions we made in this paper that
are worth discussing. First, in our characterization of the null space
we assumed we knew the stellar inclination, the stellar rotation period,
and the limb darkening coefficients
exactly, and we computed the information content of stellar light curves
in the limit of infinite SNR. None of these assumptions are valid in
practice. In realistic scenarios, the posterior shrinkage will be
necessarily \emph{lower} than what we presented here: the curves in
Figure~\ref{fig:nullspace_ensemble}, for instance, are therefore
strict upper bounds on the amount of information that can be learned about
the surface. In some cases, it may be possible to infer the inclination
from spectroscopic $v \sin I$ measurements, or the limb darkening coefficients
from the shape of transiting exoplanet light curves; but any uncertainty
in these quantities will degrade our ability to constrain
information about the surface.

Moreover, our constraints on the information content of ensembles of light curvesare
also strict upper limits, since we assumed all stars in the ensemble
are identical. Variance in the surface maps of stars in
a population makes it effectively impossible to infer detailed
properties of the individual stellar surfaces in the ensemble.
We argued without proof that what we can learn instead are properties of the
\emph{distribution} of stellar surfaces among the population, such as
the distribution of spot sizes and active latitudes.
Demonstrating this point is more difficult, so we defer it to
\citetalias{PaperII}, where we construct a custom Gaussian process model
and show that it can be used to infer such population-level parameters
from ensembles of light curves.

Finally, we limited our discussion to static stellar surfaces.
Real stellar surfaces vary
with time as individual spots form, evolve, and dissipate, or as a star progresses
through its activity cycle. This, too, makes it more difficult to constrain
the stellar surface at a particular point in time, since the amount of
data corresponding to a particular surface configuration is more limited.
On the other hand, a time-variable surface could be used to our advantage
from an ensemble standpoint. Although we are unable to uniquely infer
invididual spot properties, we could treat each point in the light curve
as a realization of a random process (i.e., a draw from some distribution)
describing the stellar surface at a high level. In this sense, we could
harness the fact that we can measure the light curves of many ``similar''
surfaces to learn something about their statistical properties.
Demonstrating this is beyond the scope of the present work, but we
return to this point in \citetalias{PaperII}.
